Exploring probability distributions for bivariate temporal granularities

### 2019-07-29 main.Rmd
adding bibliography

```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "(ref:ped-map-cap)"}
```

(ref:ped-map-cap) Google map of the Melbourne city area, dots indicate all the sensor locations. The three locations highlighted will be analyzed in the paper: the State Library is a public library; Flagstaff Station is a train station, closed on non-work days; Birrarung Marr is an outdoor park hosting many cultural and sports events.

```{r time-series-plot, fig.height = 6, fig.cap = "(ref:time-series-plot)"}
```

(ref:time-series-plot) Time series plots showing 2016 pedestrian counts, measured by three different sensors in the city of Melbourne. Small multiples of lines show that the foot traffic varies at one location to another. The spike in counts at the State Library corresponds to the timing of the event "White Night", where there were many people taking part in activities in the city throughout the night. A relatively persistent pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather noisy and spiky, with several runs of missing records.

```{r facet-time, fig.height = 6, fig.cap = "(ref:facet-time-cap)"}
```

(ref:facet-time-cap) Hourly pedestrian counts for 2016, faceted by sensors, and days of the week. The focus is on time of day and day of week across the sensors. Daily commuter patterns at Flagstaff Station, the variability of the foot traffic at Birrarung Marr, and the consistent pedestrian behavior at the State Library, can be seen.

The work is inspired by @Wickham2012glyph, which uses modulus arithmetic to display spatio-temporal data as glyphs on maps. It is also related to recent work by @R-geofacet which provides methods in the **geofacet** R package to arrange data plots into a grid, while preserving the geographical position. Both of these show data in a spatial context.

In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events. There are some existing works in this area. For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days. It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office. @Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively. However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day. Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation. Color-encoding is also low on the hierarchy of optimal variable mapping [@cleveland1984graphical; @lam2007overview].

The proposed algorithm goes beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers. The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)). Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes. An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

@@

The entire distribution can be visualized or some contextual summary statistics can be visualised to emphasize certain properties of the distribution. These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution. They can also be useful in detecting extreme behavior or anolmalies in the dataset.

hich ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.

>>>>

<!-- The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), and variations of its usage (Section \@ref(sec:variations)). Section \@ref(sec:facet-calendar) explains the full faceting extension, that is equipped with formal labels and axes. An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions. -->

@@

The entire distribution can be visualized or some contextual summary statistics can be visualised to emphasize certain properties of the distribution. These properties can throw light on central tendency, skewness, kurtosis, variation of the distribution and can also be useful in detecting extreme behavior or anolmalies in the dataset.

Which ditribution plot to draw, is based on what features of the distribution we are interested to look at, or how much display space is available to us.
